Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-10-09 paper/submitted/round3_png/algorithms_v6_cmame.tex
edits for second round of revisions

Just as error estimates can be used to guide mesh refinement~\cite{BecRann01,VendDarm00,PrudOden99,BecVex05,AinsOden11}, the error estimate obtained after dropping the remainder terms $\mathcal{R}$ and $\mathcal{R}_\mathscr{R}$ in \Cref{eq:finErrExp} can be localized to give elemental contributions and used to guide physics-based refinement over the domain to create a mixed-fidelity model. After refinement, the error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some error threshold is reached. \Cref{alg:refSeries} describes this approach. 
%Note that the QoI error adjoint problem \Cref{eq:superAdjEq} involves linearization about $\Psi_{HF}$, which is not available, so in the case of a nonlinear goal-oriented inverse problem, the QoI error adjoint problem is approximated by linearizing about $\Psi_{LF}$ instead.
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMixedFidelity}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \Cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

As an example, consider a pair of low- and high-fidelity models described by the convection-diffusion and convection-diffusion-reaction equations
\begin{subequations}
\begin{align}
\textrm{low-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u = q, \label{eq:cdvcdrLF}\\
\textrm{high-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = q, \label{eq:cdvcdrHF}
\end{align}
\end{subequations}
where $k_d$ and $k_r$ are diffusion and reaction coefficients, respectively, and the state $u$ and parameters $q$ are continuous functions over a domain $\Omega$ with homogeneous Dirichlet boundary conditions. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively; initially, $\Omega_{LF}=\Omega$. The resulting mixed-fidelity models can be described by,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= q,
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient, %that's zero in the $\Omega_{LF}$ and equal to $k_r$ in $\Omega_{HF}$; 
%
\begin{equation}
k^{MF}_r=
\begin{cases}
k_r & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Let the QoI be described by $I(q,u)=\int_{\Omega_I} u \:\textrm{d}\Omega$, where $\Omega_I\subset\Omega$ is the QoI region. Let the observations be of the state at some set of locations, so that the data-misfit term is $\frac{1}{2}\int_\Omega \delta_{obs}(u-d)^2\:\textrm{d}\Omega$, where $\delta_{obs}$ is a sum of Dirac delta functions indicating the locations of the observations. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}, with $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla q\|_2^2\:\textrm{d}\Omega$, where $\beta$ is a regularization coefficient. Assuming divergence-free velocity $\vec{V}$, the mixed-fidelity model has augmented Lagrangian
\begin{eqnarray}
\mathcal{M}_{MF}((q,u,z)(p,v,y)) &=& \int_{\Omega_I\subset\Omega} u \:\textrm{d}\Omega \nonumber\\
&+& \int_\Omega p(-\beta\nabla^2q+z) \nonumber\\
&+& v(k_d\nabla^2z+\vec{V}\cdot\nabla z+2k^{MF}_rzu+\delta_{obs}(u-d)) \nonumber\\ 
&+& y(k_d\nabla^2u-\vec{V}\cdot\nabla u+k^{MF}_ru^2+q)\:\textrm{d}\Omega.
\end{eqnarray}
Taking variations with respect to the auxiliary variables $(p,v,y)$ recovers the equations for the primary variables; taking variations with respect to the primary variables gives the equations for the auxiliary variables. Solving for the primary variables and then the auxiliary variables gives the stationary point $\Psi_{MF}$ of $\mathcal{M}_{MF}$. 
The high-fidelity versions of these primary and auxiliary systems are satisfied by $\Psi_{HF}$; we seek the error in the linear functional described by \Cref{eq:supadjout} induced by applying it to $\Psi_{MF}$ instead of $\Psi_{HF}$. We take the classical dual-weighted residual approach and solve for the supplementary adjoint $\Lambda$, linearizing about $\Psi_{MF}$ instead of $\Psi_{HF}$ when necessary, since the latter is not available; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.
%The weighted residual of the high-fidelity primary and auxiliary equations, evaluated at $\Psi_{MF}$, describes a linear functional in the weights (see \Cref{eq:supadjout}). Taking the adjoint of these high-fidelity primary and auxiliary systems, with output described by this linear functional, gives the supplementary adjoint; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena (as in the example above), or be a reduced-order model, or have a reduced parameter and/or state space. The derived error estimate is not applicable to all models, however. It is assumed that the two models share the same domain; for example, the low-fidelity model can not be a one-dimensional average of a two-dimensional high-fidelity model. The low- and high-fidelity models must be expressible in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. We need some degree of compatibility between the two models, in that we require the low-fidelity solutions to be interpretable in the context of the high-fidelity model; specifically, we require that $\Psi_{LF}$ be in a space admissible to $\mathcal{M}'_{HF}$ and $\mathcal{M}_{HF}$. For example, compared to the high-fidelity model, the low-fidelity model may have a coarser mesh, basis functions of a lower order, or basis functions spanning a reduced subspace. The high-fidelity model could describe changes in states that have a fixed nominal value in the low-fidelity model; the high-fidelity model could allow anisotropy of the inferred parameters while the low-fidelity model could constrain their components and require isotropy. We note that in the case where the two models correspond to two levels of mesh refinement, method described in~\cite{BecVex05} would likely be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead.
%, and that the QoI functional $I$ is applicable to both $(q_{HF},u_{HF})$ and $(q_{LF},u_{LF})$. %implied by admissibility of LF to HF

>>>>

Just as error estimates can be used to guide mesh refinement~\cite{BecRann01,VendDarm00,PrudOden99,BecVex05,AinsOden11}, the error estimate obtained after dropping the remainder terms $\mathcal{R}$ and $\mathcal{R}_\mathscr{R}$ in \Cref{eq:finErrExp} can be localized to give elemental contributions and used to guide physics-based refinement over the domain to create a mixed-fidelity model. After refinement, the error estimate can be calculated again, using the mixed-fidelity model as the lower-fidelity model. This process can be repeated, successively increasing the proportion of the domain in which the high-fidelity model is used, until some error threshold is reached. \Cref{alg:refSeries} describes this approach. 
%
\alglanguage{pseudocode}
\begin{algorithm}[h!]
\small
\caption{An algorithm to adaptively build a mixed-fidelity model for low error in the QoI.}
\label{alg:refSeries}
\begin{algorithmic}[1]
\State{Define maximum acceptable absolute relative QoI error \texttt{errTol}}
\State{Define maximum number of adaptive iterations \texttt{maxIter}}
\Procedure{$\texttt{BuildMixedFidelity}$}{HF model, LF model, \texttt{errTol}, \texttt{maxIter}}
	\State{Let the model MF$_0$ be the LF model applied everywhere in the domain.}
	\State{$i\gets0$}
	\State{Solve for stationary point $\Psi_{MF_0}$ of augmented Lagrangian $\mathcal{M}_{MF_0}$}
	\State{Solve QoI error adjoint equation, linearized about $\Psi_{MF_0}$, for \par\hskip\algorithmicindent supplementary adjoint $\Lambda_0$ (see \Cref{eq:superAdjEq})}
	\State{Compute QoI error estimate

@@

As an example, consider a pair of low- and high-fidelity models described by the convection-diffusion and convection-diffusion-reaction equations
\begin{subequations}
\begin{align}
\textrm{low-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u = q, \label{eq:cdvcdrLF}\\
\textrm{high-fidelity:   } & k_d\nabla^2 u - \vec{V}\cdot\nabla u + k_ru^2 = q, \label{eq:cdvcdrHF}
\end{align}
\end{subequations}
where $k_d$ and $k_r$ are diffusion and reaction coefficients, respectively, and the state $u$ and parameters $q$ are continuous functions over a domain $\Omega$ with homogeneous Dirichlet boundary conditions. To form the mixed-fidelity models, we divide the domain into complementary subdomains, $\Omega_{HF}$ and $\Omega_{LF}$, where the high- and low-fidelity models are solved, respectively; initially, $\Omega_{LF}=\Omega$. The resulting mixed-fidelity models can be described by,
%
\begin{equation}
k_d\nabla^2 u - \vec{V}\cdot\nabla u + k^{MF}_ru^2= q,
\end{equation}
%
where $k^{MF}_r$ is a piecewise-constant reaction coefficient,
%
\begin{equation}
k^{MF}_r=
\begin{cases}
k_r & \textrm{if }x\in\Omega_{HF} \\
0 & \textrm{if }x\in\Omega_{LF}.
\end{cases}
\end{equation}
%
Let the QoI be described by $I(q,u)=\int_{\Omega_I} u \:\textrm{d}\Omega$, where $\Omega_I\subset\Omega$ is the QoI region. Let the observations be of the state at some set of locations, so that the data-misfit term is $\frac{1}{2}\int_\Omega \delta_{obs}(u-d)^2\:\textrm{d}\Omega$, where $\delta_{obs}$ is a sum of Dirac delta functions indicating the locations of the observations. Since the inverse problem is ill-posed, we use Tikhonov regularization~\cite{EngHanNeu00}, with $R(q)=\frac{\beta}{2}\int_\Omega \|\nabla q\|_2^2\:\textrm{d}\Omega$, where $\beta$ is a regularization coefficient. Assuming divergence-free velocity $\vec{V}$, the mixed-fidelity model has augmented Lagrangian
\begin{eqnarray}
\mathcal{M}_{MF}((q,u,z)(p,v,y)) &=& \int_{\Omega_I\subset\Omega} u \:\textrm{d}\Omega \nonumber\\
&+& \int_\Omega p(-\beta\nabla^2q+z) \nonumber\\
&+& v(k_d\nabla^2z+\vec{V}\cdot\nabla z+2k^{MF}_rzu+\delta_{obs}(u-d)) \nonumber\\ 
&+& y(k_d\nabla^2u-\vec{V}\cdot\nabla u+k^{MF}_ru^2+q)\:\textrm{d}\Omega.
\end{eqnarray}
Taking variations with respect to the auxiliary variables $(p,v,y)$ recovers the equations for the primary variables; taking variations with respect to the primary variables gives the equations for the auxiliary variables. Solving for the primary variables and then the auxiliary variables gives the stationary point $\Psi_{MF}$ of $\mathcal{M}_{MF}$. 
The high-fidelity versions of these primary and auxiliary systems are satisfied by $\Psi_{HF}$; we seek the error in the linear functional described by \Cref{eq:supadjout} induced by applying it to $\Psi_{MF}$ instead of $\Psi_{HF}$. We take the classical dual-weighted residual approach and solve for the supplementary adjoint $\Lambda$, linearizing about $\Psi_{MF}$ instead of $\Psi_{HF}$ when necessary, since the latter is not available; the components of the supplementary adjoint corresponding to the primary and auxiliary systems can be solved separately. Having obtained $\Psi_{MF}$ and the supplementary adjoint $\Lambda$, we evaluate the error estimate (see Step \ref{alg-line:MFerrorest} of \Cref{alg:refSeries}), localize it (see \Cref{sec:errLocal}), and use the localization to guide the selection of elements currently in $\Omega_{LF}$ to include in $\Omega_{HF}$ in the next iteration. We return to this example and give results of applying \Cref{alg:refSeries} to this pair of models in \Cref{sec:cdvcdr}.

\Cref{alg:refSeries} is applicable to a large class of models. The lower-fidelity model could, for example, be a simplified model including fewer physical phenomena (as in the example above), or be a reduced-order model, or have a reduced parameter and/or state space. The derived error estimate is not applicable to all models, however. The low- and high-fidelity models must be expressible in a weak form, so this cannot be applied to, for example, a model of chemical reactions using kinetic Monte Carlo. In addition, as described in \Cref{sec:setup}, the models must be compatible, in that the low-fidelity parameters and states $q_{LF}$ and $u_{LF}$ must lie in the high-fidelity parameter and state spaces as well. This implies the two models must share the same domain; for example, the low-fidelity model can not be a one-dimensional average of a two-dimensional high-fidelity model. Models that differ in discretization can be compatible; the low-fidelity model may have a coarser mesh, basis functions of a lower order, or basis functions spanning a reduced subspace. Models that have different states and/or parameters can also be compatible; for example, the high-fidelity model could describe changes in states that have a fixed value (often zero) in the low-fidelity model. The high-fidelity model could allow anisotropy in its parameters while the low-fidelity model could constrain their components and require isotropy. We note that in the case where the two models correspond to two levels of mesh refinement, method described in~\cite{BecVex05} would likely be more efficient, since interpolation could be used to estimate $\Psi_{HF}-\Psi_{LF}$ instead.

### 2017-10-09 paper/submitted/round3_png/mathematical_formulation_v6_cmame.tex
edits for second round of revisions

A given physical system can be described to varying degrees of fidelity using different models. One has a choice in how to define the stucture of the inferred parameters $q$, as well as how to define the state equation and variables. For example, the parameter $q$ can be modeled as a single scalar value or as an infinite-dimensional distributed field variable. The state equation can be nonlinear and multiscale, requiring an expensive computation to resolve (e.g.\ the Navier-Stokes equations) or a linear, single physics model (e.g.\ Laplace's equation). All these different modeling choices lead to different versions of the optimization problem described by \Cref{eq:invOpt} that would need to be solved to infer the unknown parameters.

We assume that there exists an `ideal' model for solving the inverse problem, which we will call the high-fidelity model; given adequate computational resources, we would use this high-fidelity to formulate and solve our inverse problem. We use the subscript $_{HF}$ to denote the high-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptHF}
\begin{align}
q_{HF},u_{HF} &=&\arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objHF} \\
&&\textrm{s.t. }& \quad a_{HF}(u,q)(\phi)=\ell_{HF}(q)(\phi),\quad\forall\phi\in U_{HF} \label{eq:invOpt_consHF}
\end{align}
\end{subequations}
%
where $q_{HF} \in Q_{HF}$ and $u_{HF}\in U_{HF}$ are the inferred parameters and corresponding state, respectively. In practice, one has limited computational resources, so one may need to instead infer the parameters using a less computationally demanding model, which we call the low-fidelity model. We use the subscript $_{LF}$ to denote the low-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptLF}
\begin{align}
q_{LF},u_{LF} &=& \arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objLF} \\
&& \textrm{s.t. }& \quad a_{LF}(u,q)(\phi)=\ell_{LF}(q)(\phi),\quad\forall\phi\in U_{LF}, \label{eq:invOpt_consLF}
\end{align}
\end{subequations}
%
where $q_{LF} \in Q_{LF}$ and $u_{LF}\in U_{LF}$ are the inferred parameters and corresponding state, respectively. If there is compatibility between the low-fidelity and high-fidelity solution spaces for the parameter and state variables (i.e., if the solutions of one can be interpreted in the context of the other; see \Cref{sec:adapt_alg} for more details), one may combine the two models to form a mixed-fidelity model and corresponding inverse problem. The low-fidelity model can be used in part of the computational domain, and the high fidelity in the remainder; we denote the mixed-fidelity model and its variables with the subscript $_{MF}$. How one chooses to combine the low- and high-fidelity models can be informed by what aspect of the inferred parameters is of interest.

@@

%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, we take the QoI calculated from inferring the parameters with the highest-fidelity model to be the `exact' or `true' solution; we now derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model.
%
\begin{proposition}
\label{thm:error_estimate}
Consider the high- and low-fidelity inverse problems described by the constrained optimization problems in \Cref{eq:invOptHF,eq:invOptLF}. Let the forms $a_{HF/LF}:U_{HF/LF} \times Q_{HF/LF} \times U_{HF/LF} \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U_{HF/LF}\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q_{HF/LF}\to\Reals$ be differentiable with respect to the parameter $q$, and the Quantity of Interest (QoI) functional $I:Q_{HF/LF}\times U_{HF/LF}\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

@@

%
\begin{proposition}
\label{thm:error_estimate_dual}
Let the hypotheses of Proposition~\ref{thm:error_estimate} hold. Let the variational form
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys_prop}
\end{equation}
%
represent the equation $\mathcal{M}'_{HF,\Psi}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define an adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
for the supplementary adjoint $\Lambda$, where $\mathscr{R}$ is defined in \Cref{eq:supadjsys_prop}, and where
%Let $\Lambda$ be the solution of the dual problem, \red{$\mathcal{M}'$ has just one '?}
%
%\begin{equation}
%\mathcal{M}_{HF}'(\Psi_{LF};\Phi, \Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
%\end{equation}
%
where $\mathcal{Q}:(U_{HF} \times U_{HF} \times Q_{HF})^2 \to \R$ is a linear functional defined by,
%
\begin{equation}
\mathcal{Q}(\Phi)=\mathcal{M}'_{HF}(\Psi_{LF})(\Phi).
\label{eq:supadjout}
\end{equation}
%
Then we have,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF}) + \mathcal{R}_{\mathscr{R}} + \mathcal{R}
\end{multline}
%
where $\mathcal{R}_{\mathscr{R}}$ is a second-order term in the error $e=\Psi_{HF}-\Psi_{LF}$, and $\mathcal{R}$ is as defined in \Cref{thm:error_estimate}. %\red{Now that we've explicitly pointed out the second-order error term, can we really say that we derived "an adjoint-based third-order estimate" (in abstract and intro)?}
%
\end{proposition}
%
\begin{proof}
%
The proof simply follows from the classic dual-based a posteriori error estimation strategy. Note that, as a stationary point, $\Psi_{HF}$ satisfies,
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys}
\end{equation}
%
which is represents the equation $\mathcal{M}'_{HF}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define the adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
%
for the supplementary adjoint $\Lambda$. The error in the linear output $\mathcal{Q}$ defined in \Cref{eq:supadjout} can then be expressed as a dual-weighted residual,
%
\begin{equation}
\label{eq:adjOutErr}
\mathcal{Q}(\Psi_{HF}-\Psi_{LF})=-\mathscr{R}(\Psi_{LF})(\Lambda) + R_{\mathscr{R}},
\end{equation}
%
where $R_{\mathscr{R}}$ is a second order remainder term due to the linearization of $\mathscr{R}$ about $\Psi_{LF}$ in \Cref{eq:superAdjEq}. This implies the required result. %\red{What is the classical notation for adjoint problems? Is it okay that we're breaking from the previous notation where a second pair of parentheses denotes things that we are linear in?}
%
\end{proof}
%
%We note that if the state equations of the high-fidelity model are linear, the remainder term $R_\mathscr{R}$ in the supplementary adjoint formulation in \Cref{eq:adjOutErr} vanishes, and the error estimate in \Cref{eq:finErrExp} becomes fully third-order. This occurs even if the QoI functional $I$ is nonlinear.
%If the state equations of the high fidelity model are linear, the derivative of the Lagrangian in~\Cref{eq:InvsOpt_auglag} is linear. In this case, the $R_{\mathscr{R}}$ term in~\Cref{eq:finErrExp} disappears, even if the QoI functional $I$ is nonlinear, and we get a fully third order error estimate.

>>>>

A given physical system can be described to varying degrees of fidelity using different models. One has a choice in how to define the structure of the inferred parameters $q$, as well as how to define the state equation and variables. For example, the parameter $q$ can be modeled as a single scalar value or as a distributed field variable. The state equation can be nonlinear and multiscale, requiring an expensive computation to resolve (e.g.\ the Navier-Stokes equations) or a linear, single physics model (e.g.\ Laplace's equation). All these different modeling choices lead to different versions of the optimization problem described by \Cref{eq:invOpt} that would need to be solved to infer the unknown parameters. Note that in order to solve \Cref{eq:invOpt} numerically, the description of a model must usually include a choice of discretization method in addition to a choice of equations.

We assume that there exists an `ideal' model for solving the inverse problem, which we will call the high-fidelity model; given adequate computational resources, we would use this high-fidelity to formulate and solve our inverse problem. We use the subscript $_{HF}$ to denote the high-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptHF}
\begin{align}
q_{HF},u_{HF} &=&\arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objHF} \\
&&\textrm{s.t. }& \quad a_{HF}(u,q)(\phi)=\ell_{HF}(q)(\phi),\quad\forall\phi\in U_{HF} \label{eq:invOpt_consHF}
\end{align}
\end{subequations}
%
where $q_{HF} \in Q_{HF}$ and $u_{HF}\in U_{HF}$ are the inferred parameters and corresponding state, respectively. In practice, one has limited computational resources, so one may need to instead infer the parameters using a less computationally demanding model, which we call the low-fidelity model. We use the subscript $_{LF}$ to denote the low-fidelity inverse problem,
%
\begin{subequations}
\label{eq:invOptLF}
\begin{align}
q_{LF},u_{LF} &=& \arg\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_objLF} \\
&& \textrm{s.t. }& \quad a_{LF}(u,q)(\phi)=\ell_{LF}(q)(\phi),\quad\forall\phi\in U_{LF}, \label{eq:invOpt_consLF}
\end{align}
\end{subequations}
%
where $q_{LF} \in Q_{LF}$ and $u_{LF}\in U_{LF}$ are the inferred parameters and corresponding state, respectively.

Assume there is compatibility between the low-fidelity and high-fidelity parameter and state solution spaces; specifically, assume that the low-fidelity parameters and states $q_{LF}$ and $u_{LF}$ lie in the high-fidelity parameter and state spaces as well, so that we have $q_{LF}\in Q_{HF}$ and $u_{LF}\in U_{HF}$ (see \Cref{sec:adapt_alg} for interpretation details). Given this compatibility, one may combine the two models to form a mixed-fidelity model and corresponding mixed-fidelity inverse problem. The low-fidelity model can be used in part of the computational domain, and the high fidelity in the remainder; we denote the mixed-fidelity model and its variables with the subscript $_{MF}$. How one chooses to combine the low- and high-fidelity models can be informed by what aspect of the inferred parameters is of interest.

@@

%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of compatible models, we take the QoI calculated from inferring the parameters with the highest-fidelity model to be the `exact' or `true' solution; we now derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model.
%
\begin{proposition}
\label{thm:error_estimate}
Consider the high- and low-fidelity inverse problems described by the constrained optimization problems in \Cref{eq:invOptHF,eq:invOptLF}. Let the forms $a_{HF/LF}:U_{HF/LF} \times Q_{HF/LF} \times U_{HF/LF} \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U_{HF/LF}\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q_{HF/LF}\to\Reals$ be differentiable with respect to the parameter $q$, and the QoI functional $I:Q_{HF/LF}\times U_{HF/LF}\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

@@

%
\begin{proposition}
\label{thm:error_estimate_dual}
Let the hypotheses of Proposition~\ref{thm:error_estimate} hold. Let the variational form
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys_prop}
\end{equation}
%
represent the equation $\mathcal{M}'_{HF,\Psi}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define an adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
for the supplementary adjoint $\Lambda$, where $\mathscr{R}$ is defined in \Cref{eq:supadjsys_prop}, and where
%Let $\Lambda$ be the solution of the dual problem, \red{$\mathcal{M}'$ has just one '?}
%
%\begin{equation}
%\mathcal{M}_{HF}'(\Psi_{LF};\Phi, \Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
%\end{equation}
%
where $\mathcal{Q}:(U_{HF} \times U_{HF} \times Q_{HF})^2 \to \R$ is a linear functional defined by,
%
\begin{equation}
\mathcal{Q}(\Phi)=\mathcal{M}'_{HF}(\Psi_{LF})(\Phi).
\label{eq:supadjout}
\end{equation}
%
Then we have,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF}) + \mathcal{R}_{\mathscr{R}} + \mathcal{R}
\end{multline}
%
where $\mathcal{R}_{\mathscr{R}}$ is a second-order term in the error $e=\Psi_{HF}-\Psi_{LF}$, and $\mathcal{R}$ is as defined in \Cref{thm:error_estimate}. %\red{Now that we've explicitly pointed out the second-order error term, can we really say that we derived "an adjoint-based third-order estimate" (in abstract and intro)?}
%
\end{proposition}
%
\begin{proof}
%
The proof follows from the classic dual-based a posteriori error estimation strategy. Note that, as a stationary point, $\Psi_{HF}$ satisfies,
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys}
\end{equation}
%
which is represents the equation $\mathcal{M}'_{HF}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define the adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
%
for the supplementary adjoint $\Lambda$. The error in the linear output $\mathcal{Q}$ defined in \Cref{eq:supadjout} can then be expressed as a dual-weighted residual,
%
\begin{equation}
\label{eq:adjOutErr}
\mathcal{Q}(\Psi_{HF}-\Psi_{LF})=-\mathscr{R}(\Psi_{LF})(\Lambda) + R_{\mathscr{R}},
\end{equation}
%
where $R_{\mathscr{R}}$ is a second order remainder term due to the linearization of $\mathscr{R}$ about $\Psi_{LF}$ in \Cref{eq:superAdjEq}. This implies the required result. %\red{What is the classical notation for adjoint problems? Is it okay that we're breaking from the previous notation where a second pair of parentheses denotes things that we are linear in?}
%
\end{proof}
%
%We note that if the state equations of the high-fidelity model are linear, the remainder term $R_\mathscr{R}$ in the supplementary adjoint formulation in \Cref{eq:adjOutErr} vanishes, and the error estimate in \Cref{eq:finErrExp} becomes fully third-order. This occurs even if the QoI functional $I$ is nonlinear.
%If the state equations of the high fidelity model are linear, the derivative of the Lagrangian in~\Cref{eq:InvsOpt_auglag} is linear. In this case, the $R_{\mathscr{R}}$ term in~\Cref{eq:finErrExp} disappears, even if the QoI functional $I$ is nonlinear, and we get a fully third order error estimate.
